---
title: "GeneticVariation_Multiplicity_Methods"
output: html_document
---

#### **Overview of *easyRpopgen* methodology**

latent genetic potential (LGP) and hypothetical gametic multilocus diversity (vgam) are calculated using functions written for the application, using *poppr* functionalities.

Below is a simplified code displaying the logic behind the application calculations of each statistic and analysis conducted. You will find if you adapt this code for your own research much can be done simpler, though this set-up was used to ensure that data format was consistent and could be easily piped in to the summary table.

#### **Population Size**

```{r, eval = FALSE}
  N<-summary(genind) #summarize genind object with adegenet functions
  N<-as.data.frame((N$n.by.pop)) #Separate out N by pop as a dataframe
  N$pop<-rownames(N) #create population column
  colnames(N)<-c("N", "pop") #rename columns for neatness
  N$minN<-min(N$N) #Find smallest population size
  N #View population sizes
```

#### **Number of Alleles (A)**

```{r, eval = FALSE}
#Function to determine number of alleles for a population
Na=function (x, population)
{
  sub=poppr::popsub(x, population, drop=TRUE) #subset population
  sum=summary(sub) #summarize genind object with adegenet functions
  sum$loc.n.all #extract number of alleles per loci across individuals from summary
}

Nallele=NULL; #Create empty object to put outputs in
  for (i in unique(genind$pop)) #Loop to calculate A for each locus per population
  {
    tmp=Na(genind, population=i) #Carry out Na function
    Nallele=rbind(Nallele,tmp) #Merge results per population in Nallele
  }
  Nallele=round(as.data.frame(rowMeans(Nallele)),3) #Round A to 3 dp
  rownames(Nallele)=unique(genind$pop) #rename rows for population
  names(Nallele)="A" #rename column value name to reflect metric
  Nallele$pop=rownames(Nallele) #Add population column
  Nallele #View A per column
```

#### **Polymorphic Loci (%P)**

```{r, eval = FALSE}
  mon<-NULL;
  pol<-NULL;
  per<-NULL; #Create empty object to put outputs in
  for (i in unique(genind$pop)) { #loop to calculate monomorphic loci, polymorphic loci and percentage of polymorphic loci per loci for each pop
    temp<-poppr::popsub(genind, sublist =i, drop= TRUE) #subset population i
    temp<-summary(temp) #summarize genind object
    loc<-temp$loc.n.all #Extract locus allele data
    total<-length(temp$loc.n.all) #find total number of loci
    
    monomorphs<-as.matrix((table(loc) [names(table(loc)) == 1]))
    if(is.na(monomorphs[1]) == TRUE) {
      monomorphs = 0
    } #Find monomorphic loci
    polymorphs<-as.matrix(total - monomorphs) #calculate polymorphic loci
    P<-as.matrix((polymorphs/total) *100) #Calculate percentage of polymorphic loci
    mon<-rbind(mon, monomorphs) #Merge results per population
    pol<-rbind(pol, polymorphs) #Merge results per population
    per<-rbind(per, P) #Merge results per population
  }
  
  pop<-as.data.frame(unique(genind$pop))#Create df from pops
  FinalP<-cbind(pop, total,mon, pol, per) #Merge locus summary data with pops
  colnames(FinalP)<-c("pop", "total loci", "monomorphic loci", "polymorphic loci", "%P") #rename columns for neatness
  rownames(FinalP)<-FinalP$pop #change rownames for niceness
  FinalP$`%P`<-round(FinalP$`%P`,0) #round percentages to zero dp
  FinalP
```

#### **Allelic Richness (Ar) and Private Allelic Richness (pAr)**

The functions used to calculate allelic and private allelic richness come not from a published R package, but from those used by [Dupuis *et al* (2018)](https://link.springer.com/article/10.1007/s10592-018-1081-8) based on the approach of [Kalinowski (2004)](https://link.springer.com/article/10.1023/B:COGE.0000041021.91777.1a). The latter paper and approach also forms the basis of the non-R based program *HP-Rare* developed by the author [(Kalinowski, 2005)](https://doi.org/10.1111/j.1471-8286.2004.00845.x). The necessary function scripts to carry out the below analysis can be found in the "App" code for this application, and at the [GitHub directory](https://github.com/muirheadk/GBS_analysis_pipeline ) associated with the paper. If you use any measures of rarefied allelic richness make sure that you cite Dupuis *et al* (2018) in the same manner that you would cite a package or program.

```{r, eval = FALSE}
# Extract the data to use
  data.list = list()
  data.list$tab <- genind@tab
  data.list$loc.fac <- factor(genind@loc.fac)
  data.list$pop <- factor(genind@pop)
  
  # Run rarefaction analyses, using smallest population as minimal cutoff
  min.size <- min(table(data.list$pop))
  rarefied.values <- rarefiedMatrices(data = data.list, 
                                      g = min.size, 
                                      display.progress = TRUE)
  rarefied.values

#################################
########Allelic Richness#########
#################################
  
  richness.matrix<-rarefied.values$richness #subset Ar values
  richness.matrix<-as.data.frame(richness.matrix) #Convert object to df
  
  richness.matrix=tidyr::gather(richness.matrix, pop, Ar, na.rm = TRUE, convert = FALSE) #Rearrange data
  
  finalAr<-richness.matrix %>% dplyr::group_by(pop) %>% #Summarize richness values per population
    dplyr::summarize(
      N = length(Ar) - sum(is.na(Ar)),
      mean = mean(Ar, na.rm = TRUE),
      sd   = sd(Ar, na.rm =TRUE),
      se   = sd / sqrt(N),
    )
  
  finalAr$mean<-round(finalAr$mean, 3)
  finalAr$sd<-round(finalAr$sd, 3)
  finalAr$se<-round(finalAr$se, 3)
  names(finalAr)=c("pop","Nmin", "Ar","SD","SE")
  finalAr
  
#################################
####Private Allelic Richness#####
#################################
  
  richness.matrix<-rarefied.values$private #subset pAr values
  richness.matrix<-as.data.frame(richness.matrix) #Convert object to df
  
  richness.matrix=tidyr::gather(richness.matrix, pop, pAr, na.rm = TRUE, convert = FALSE) #Rearrange data
  
  finalpAr<-richness.matrix %>% dplyr::group_by(pop) %>% #Summarize richness values per population
    dplyr::summarize(
      N = length(pAr) - sum(is.na(pAr)),
      mean = mean(pAr, na.rm = TRUE),
      sd   = sd(pAr, na.rm =TRUE),
      se   = sd / sqrt(N),
    )
  
  finalpAr$mean<-round(finalpAr$mean, 3)
  finalpAr$sd<-round(finalpAr$sd, 3)
  finalpAr$se<-round(finalpAr$se, 3)
  names(finalpAr)=c("pop","Nmin", "pAr","SD","SE")
  finalpAr
```

#### **Latent Genetic Potential (LGP)**

LGP is measured as the sum of the number of alleles minus the effective number of alleles ($LGP = \sum_i^I Ai - Ae_i$ where i is a given locus and I is all loci considered). Due to this it is therefore important to consider that it is a value which is incredibly sensitive to the number of loci considered and not necessarily comparable between studies because of this.

```{r, eval = FALSE}
#LGP calculation function
LGPcalc=function (x, population = "ALL")
{
  z = poppr::popsub(x, population, drop = TRUE) #seperates genind by population
  sum=summary(z)  #generates summary table of genind object including measures of 'number of alleles' and 'expected heterozygosity' which are needed to calculate LGP
  q=(1/(1-sum$Hexp))  #Calculate number of effective alleles
  LGP=(sum(sum$loc.n.all-q)) #calculate LGP
}

#Calculation
LGP=NULL; #Create empty object to put outputs in
  for (i in unique(genind$pop)) #Loop to calculate LGP across populations
  {
    temp<-LGPcalc(genind, population = i) #Carry out LGP calculation for pop i
    LGP=rbind(LGP,temp) #Merge LGP results from all pops
  }
  LGP<-as.data.frame(LGP) #Convert object to df
  rownames(LGP)=unique(genind$pop) #Add pops from genind as rownames
  names(LGP)="LGP" #rename value for neatness
  LGP$pop=rownames(LGP) #Create pop column
  LGP #View LGP values per population
```

#### **Hypothetical Gametic Multilocus Diversity (vgam)**

vgam is the product of the expected number of alleles across loci within a population. Like LGP the value is therefore incredibly sensitive to the number of loci considered, and can therefore appear unweildingly large when considering large SNP datasets. Consider this when interpreting the results, and note that due to this it is also not necessarily comparable between studies.

```{r, eval = FALSE}
#vgam calculation function
#vgam
Vgamcalc=function (x, population = "ALL") #Function to calculate vgam from a genind object
{
  z=poppr::popsub(x, sublist = population, drop = TRUE) #subset genind by population
  sum=summary(z) #generate summary statistics including expected heterozygosity
  q=as.matrix(1/(1-sum$Hexp)) #calculate effective number of alleles from expected heterozygosity
  vgam=colProds(q,method = c("direct")) #calculate hypothetical gametic multi-locus diversity as product of all Aes across all loci
  vgam
}

#Calculation
vgam=NULL; #Create empty object to put outputs in
for (i in unique(genind$pop)) #loop to calculate vgam per population
{
  temp<-Vgamcalc(genind, population = i) #calculate vgam for population i
  vgam=rbind(vgam,temp) #merge values of all populations together
}
vgam<-as.data.frame(vgam) #convert object to df
rownames(vgam)=unique(genind$pop) #rename rows for populations from genind object
names(vgam)="vgam" #rename value column for neatness
vgam$pop=rownames(vgam) #create pop column
vgam #view vgam per population
```